A 3D curvilinear skeletonization algorithm with application to path tracing

### 2012-12-03 skeleton_and_pathtracing.tex
Corrections integrees

\def\mydist1{D}
\def\mydd1{\Omega}
\def\mydecent{\mathcal{DC}}
\def\dist1{d}
\def\Dist1#1{\mydist1(#1)}
\def\DD1#1{\mydd1(#1)}
\def\decent#1{\mydecent(#1)}
\def\cplDist1#1{\mydist1^{cc}(#1)}
\def\cplDD1#1{\mydd1^{cc}(#1)}
\def\cpldecent#1{\mydecent^{cc}(#1)}

@@

\def\V2C{\Psi}
\def\Nset{\mathbb{N}}
\def\Zset{\Z}
\def\Nset{\mathbb{N}}
\def\Rset{\mathbb{R}}
\def\Z{\mathbb{Z}}
\def\Zhalf{\frac{\mathbb{Z}}{2}}
\def\allfaces#1{\mathbb{F}^{#1}}
\def\subfaces#1#2{\allfaces{#1}_{#2}}
\def\subcomplex#1#2{#1 \preceq #2}
\def\complex#1#2{\subcomplex{#2}{\allfaces{#1}}}
\def\cell#1{\hat{#1}}
\def\cellstrict#1{\hat{#1}^*}
\def\closure#1{#1^-}
\def\princ#1{#1^+}
\def\mydim#1{\dim(#1)}
\def\type#1{\mbox{\it{}Dir\/}(#1)}
\def\orient#1{\mbox{\it{}Orient\/}(#1)}
\def\border#1{\mbox{\it{}Border\/}(#1)}
\def\suchthat{\;|\;}
\def\inter{\cap}
\def\union{\cup}
\def\dinterval#1#2{\{#1,\ldots,#2\}}
\newcommand{\Compl}[1]{\overline{#1}}
\def\finproof{\square}
\def\Zhalf{\frac{\mathbb{Z}}{2}}
\newcommand{\Card}[1]{|#1|}

@@

We denote by $\allfaces{n}$ the set composed of all faces in $\Z^n$. Given $m \in \dinterval{0}{n}$, we denote by $\subfaces{n}{m}$ the set composed of all $m$-faces in $\Z^n$.
Let $f \in \allfaces{n}$. We set $\cell{f} = \{g \in \allfaces{n}\vert g \subseteq f\}$, and $\cellstrict{f} = \cell{f} \setminus \{f\}$. Any element of $\cell{f}$ (resp. $\cellstrict{f}$) is \myem{a face of $f$} (resp. \myem{a proper face of $f$}).
The \myem{closure} of a set of faces $X$ is the set $\closure{X}=\union\{\cell{f} \vert f \in X\}$.

@@

\begin{algorithm}[tb]
\label{algo::pardircollapse_v0}
\SetKwFor{ForAll}{for all}{do}{end}
\caption[ParDirCollapse]{$ParDirCollapse(X,W,\ell)$}
\KwData{A cubical complex $\complex{n}{X}$, a subcomplex
  $\subcomplex{W}{X}$ which represents faces of $X$ which should not
  be removed, and $\ell \in \Nset$, the number of layers of free faces
  which should be removed from $X$} 
\KwResult{A cubical complex}
\lnl{pc:01}\While{there exists free faces in $X \setminus W$ and $\ell > 0$}{
\lnl{pc:02}  $L=\border{X}^-$\;
\lnl{pc:03}  \For{$t=0 \rightarrow n-1$}{
\lnl{pc:04}    \For{$s=0 \rightarrow 1$}{
\lnl{pc:05}      \For{$d=n \rightarrow 1$}{
\lnl{pc:06}        $E=\{(f,g)$ free for $X \suchthat g \notin W,$\\
                      $\;$~~~~~~~$\type{f,g}=t,$ $\orient{f,g}=s,$ $\mydim{f}=d\}$\;
\lnl{pc:07}        $G=\{(f,g) \in E \suchthat f \in L$ and $g \in L\}$\;
\lnl{pc:08}        $X=X \setminus G$\;
                 }
               }
             }
\lnl{pc:09}  $l = l-1$\;
	   }
\lnl{pc:10} \Return $X$\;
\end{algorithm}

@@

In the following, we introduce a new method in the cubical complex framework, requiring no user input, for obtaining a curvilinear skeleton yielding satisfactory geometrical properties. Our method is based on the two previously listed strategies: it finds, during thinning, elements whose neighbourhood configuration is interesting, and uses a function on these elements to decide whether to preserve them, or not, in the result.

@@

\begin{definition}
Let $f \in \complex{n}{X}$, the \myem{death date of $f$ in $X$}, denoted by $\death{f}{X}$, is the smallest integer $d$ such that $f \notin ParDirCollapse(X, \emptyset, d)$. 
\end{definition}

@@

\begin{definition}
Let $f \in \complex{n}{X}$, the \myem{birth date of $f$ in $X$}, denoted by $\birth{f}{X}$, is the smallest integer $b$ such that either $f$ is a facet of $ParDirCollapse(X, \emptyset, b)$, or $f \notin ParDirCollapse(X, \emptyset, b)$.
\end{definition}

@@

\begin{definition}
Let $f \in \complex{n}{X}$, the lifespan of $f$ in $X$ is the integer

@@

Two points $x,y \in \Zset^n$ are \myem{$1$-neighbours} if the Euclidean distance between $x$ and $y$ is equal or inferior to $1$ (also called direct neighbours). A \myem{$1$-path from $x$ to $y$} is a sequence $\mathcal{C}=(z_0,...,z_k)$ of points of $\Zset^n$ such that $z_0=x$, $z_k=y$, and for all $j \in [1;k]$, $z_j$ and $z_{j-1}$ are $1$-neighbours. The length of $\mathcal{C}$ is $k$.

We set $\dist1(x,y)$ as the length of a shortest $1$-path from $x$ to $y$; remark that $\dist1$ is indeed the $6$-distance in the 3D voxel framework. Let $S \subset \Zset^n$, we set $\Compl{S} = \Zset^n \setminus S$, and for all $x \in \Zset^n$, the map $\Dist1{S}: \Zset^n \rightarrow \Nset$ is such that $\Dist1{S}(x) = \displaystyle \min_{y \in \Compl{S}} \dist1(x,y)$.

The \myem{maximal $1$-ball of $S$ centered on $x$} is the set $\MaxOneBall{S}(x)=\{y \in \Zset^n | \dist1(x,y) < \Dist1{S}(x)\}$.
We set, for all $x \in S$, the map $\DD1{S}: \Zset^n \rightarrow \Nset$ such that $\DD1{S}(x) = \displaystyle \max_{x \in \MaxOneBall{S}(y)} \Dist1{S}(y)$: this value indicates the radius of a largest maximal $1$-ball contained in $S$ and containing $x$. If $x \in \Compl{S}$, we set $\DD1{S}(x) = 0$. The map $\DD1{S}$ is known as the opening function of $S$ based on the $1$-distance (also called the granulometry function) \cite{Matheron67}: it allows to compute efficiently results of morphological openings by balls of various radius, and gives information on the local thickness of an object.

@@

\begin{figure}[tb]
\begin{center}
\begin{tabular}{cccc}
	\myincludegraphics{0.22\textwidth}{fish.pdf} &
	\myincludegraphics{0.22\textwidth}{fishd.pdf} &
	\myincludegraphics{0.22\textwidth}{fishdd.pdf} &
	\myincludegraphics{0.22\textwidth}{fishcenter.pdf} \\
\end{tabular}
\caption[Example of opening and decenterness map]{\label{fig::globalopening} \textbf{Example of opening and decenterness map - } From left to right: a shape $S \subset \Zset^2$ (in gray), $\Dist1{S}$, $\DD1{S}$ and $\decent{S}$ (low values have dark colour).}
\end{center}
\end{figure}

@@

\begin{algorithm}[tb]
\label{algo::3d_curvskel}
\caption[CurvilinearSkeleton]{$CurvilinearSkeleton(X)$}
\KwData{A cubical complex $\complex{3}{X}$}
\KwResult{A cubical complex $\complex{3}{Y}$}
\lnl{curvskel:01} $W=\{f \in X | \lifespan{f}{X} > \cpldecent{X}(f) + \birth{f}{X} - \cplDist1{X}(f)$ and $\mydim{f}=1\}$\;
\Return $ParDirCollapse(X,W,+\infty)$\;
\end{algorithm}

@@

>>>>

\def\mydist1{D_1}
\def\mydd1{\Omega_1}
\def\mydecent{\mathcal{DC}_1}
\def\dist1{d_1}
\def\Dist1#1{\mydist1(#1)}
\def\DD1#1{\mydd1(#1)}
\def\decent#1{\mydecent(#1)}
\def\cplDist1#1{\mydist1^{cc}(#1)}
\def\cplDD1#1{\mydd1^{cc}(#1)}
\def\cpldecent#1{\mydecent^{cc}(#1)}

@@

\def\V2C{\Psi}
\def\Nset{\mathbb{N}}
\def\Zset{\Z}
\def\Nset{\mathbb{N}}
\def\Rset{\mathbb{R}}
\def\Z{\mathbb{Z}}
\def\Zhalf{\frac{\mathbb{Z}}{2}}
\def\allfaces#1{\mathbb{F}^{#1}}
\def\subfaces#1#2{\allfaces{#1}_{#2}}
\def\subcomplex#1#2{#1 \preceq #2}
\def\complex#1#2{\subcomplex{#2}{\allfaces{#1}}}
\def\cell#1{\hat{#1}}
\def\cellstrict#1{\hat{#1}^*}
\def\closure#1{\widehat{#1}}
\def\princ#1{#1^+}
\def\mydim#1{\dim(#1)}
\def\type#1{\mbox{\it{}Dir\/}(#1)}
\def\orient#1{\mbox{\it{}Orient\/}(#1)}
\def\border#1{\mbox{\it{}Border\/}(#1)}
\def\suchthat{\;|\;}
\def\inter{\cap}
\def\union{\cup}
\def\dinterval#1#2{\{#1,\ldots,#2\}}
\newcommand{\Compl}[1]{#1^c}
\def\finproof{\square}
\def\Zhalf{\frac{\mathbb{Z}}{2}}
\newcommand{\Card}[1]{|#1|}

@@

Given $m \in \dinterval{0}{n}$, we denote by $\subfaces{n}{m}$ the set composed of all $m$-faces in $\Z^n$. We denote by $\allfaces{n}$ the set composed of all $m$-faces in $\Z^n$ : $\allfaces{n} = \bigcup \limits_{m \in [0;n]} \subfaces{n}{m}$.
Let $f \in \allfaces{n}$. We set $\cell{f} = \{g \in \allfaces{n}\vert g \subseteq f\}$, and $\cellstrict{f} = \cell{f} \setminus \{f\}$. Any element of $\cell{f}$ (resp. $\cellstrict{f}$) is \myem{a face of $f$} (resp. \myem{a proper face of $f$}).
The \myem{closure} of a set of faces $X$ is the set $\closure{X}=\union\{\cell{f} \vert f \in X\}$.

@@

\begin{algorithm}[tb]
\label{algo::pardircollapse_v0}
\SetKwFor{ForAll}{for all}{do}{end}
\caption[ParDirCollapse]{$ParDirCollapse(X,W,\ell)$}
\KwData{A cubical complex $\complex{n}{X}$, a subcomplex
  $\subcomplex{W}{X}$ which represents faces of $X$ which should not
  be removed, and $\ell \in \Nset$, the number of layers of free faces
  which should be removed from $X$} 
\KwResult{A cubical complex}
\While{there exists free faces in $X \setminus W$ and $\ell > 0$}{  \label{pc:01}
$L=\closure{\border{X}}$\; \label{pc:02}  
\For{$t=1 \rightarrow n$}{ \label{pc:03}  
\For{$s=0 \rightarrow 1$}{ \label{pc:04}  
\For{$d=n \rightarrow 1$}{ \label{pc:05}
$E=\{(f,g)$ free for $X \suchthat g \notin W,$
                      $\type{f,g}=t,$ $\orient{f,g}=s,$ $\mydim{f}=d\}$\;  \label{pc:06}       
$G=\{(f,g) \in E \suchthat f \in L$ and $g \in L\}$\; \label{pc:07}  
$X=X \setminus G$\;  \label{pc:08}
}
}
}
$l = l-1$\; \label{pc:10}
}
\Return $X$\; \label{pc:11}
\end{algorithm}

@@

In the following, we introduce a new method in the cubical complex framework, requiring no user input, for obtaining a curvilinear skeleton yielding satisfactory geometrical properties. Our method is based on the two previously listed strategies: it finds, during thinning, elements with a specific neighbourhood configuration, and uses a function on these elements to decide whether to preserve them, or not, in the result.

@@

\begin{definition}
Let $f \in \complex{n}{X}$. The \myem{death date of $f$ in $X$}, denoted by $\death{f}{X}$, is the smallest integer $\delta$ such that $f \notin ParDirCollapse(X, \emptyset, \delta)$. 
\end{definition}

@@

\begin{definition}
Let $f \in \complex{n}{X}$. The \myem{birth date of $f$ in $X$}, denoted by $\birth{f}{X}$, is the smallest integer $b$ such that either $f$ is a facet of $ParDirCollapse(X, \emptyset, b)$, or $f \notin ParDirCollapse(X, \emptyset, b)$.
\end{definition}

@@

\begin{definition}
Let $f \in \complex{n}{X}$. The lifespan of $f$ in $X$ is the integer

@@

%Two points $x,y \in \Zset^n$ are \myem{$1$-neighbours} if the Euclidean distance between $x$ and $y$ is equal to $1$ (also called direct neighbours). A \myem{$1$-path from $x$ to $y$} is a sequence $\mathcal{C}=(z_0,...,z_k)$ of points of $\Zset^n$ such that $z_0=x$, $z_k=y$, and for all $j \in [1;k]$, $z_j$ and $z_{j-1}$ are $1$-neighbours. The length of $\mathcal{C}$ is $k$.

%We set $\dist1(x,y)$ as the length of a shortest $1$-path from $x$ to $y$; remark that $\dist1$ is indeed the $6$-distance in the 3D voxel framework.

We set $\dist1(x,y)$ as the L1 distance between $x$ and $y$ (Manhattan distance). Let $S \subset \Zset^n$, we set $\Compl{S} = \Zset^n \setminus S$, and for all $x \in \Zset^n$, the map $\Dist1{S}: \Zset^n \rightarrow \Nset$ is such that $\Dist1{S}(x) = \displaystyle \min_{y \in \Compl{S}} \dist1(x,y)$.

The \myem{maximal $1$-ball in $S$ centered on $x$} is the set $\MaxOneBall{S}(x)=\{y \in \Zset^n | \dist1(x,y) < \Dist1{S}(x)\}$.
We set, for all $x \in S$, the map $\DD1{S}: \Zset^n \rightarrow \Nset$ such that $\DD1{S}(x) = \displaystyle \max_{x \in \MaxOneBall{S}(y)} \Dist1{S}(y)$: this value indicates the radius of a largest maximal $1$-ball contained in $S$ and containing $x$. If $x \in \Compl{S}$, we set $\DD1{S}(x) = 0$. The map $\DD1{S}$ is known as the opening function of $S$ based on the $1$-distance (also called the granulometry function) \cite{Matheron67}: it allows to compute efficiently results of morphological openings by balls of various radius, and gives information on the local thickness of an object.

@@

\begin{figure}[tb]
\begin{center}
\begin{tabular}{cccc}
	\myincludegraphics{0.22\textwidth}{fish.pdf} &
	\myincludegraphics{0.22\textwidth}{fishd.pdf} &
	\myincludegraphics{0.22\textwidth}{fishdd.pdf} &
	\myincludegraphics{0.22\textwidth}{fishcenter.pdf} \\
\end{tabular}
\caption[Examples of opening and decenterness map]{\label{fig::globalopening} \textbf{Examples of opening and decenterness map - } From left to right: a shape $S \subset \Zset^2$ (in gray), $\Dist1{S}$, $\DD1{S}$ and $\decent{S}$ (low values have dark colour).}
\end{center}
\end{figure}

@@

\begin{algorithm}[tb]
\label{algo::3d_curvskel}
\caption[CurvilinearSkeleton]{$CurvilinearSkeleton(X)$}
\KwData{A cubical complex $\complex{3}{X}$}
\KwResult{A cubical complex $\complex{3}{Y}$}
$W=\{f \in X | \lifespan{f}{X} > \cpldecent{X}(f) + \birth{f}{X} - \cplDist1{X}(f)$ and $\mydim{f}=1\}$  \label{curvskel:01}  \; 
\Return $ParDirCollapse(X,W,+\infty)$\;
\end{algorithm}

@@

Although the output of Alg.\ref{algo::3d_curvskel} may contain $2$-faces (depending on the topological properties of the input), the algorithm is said to be a \myem{curvilinear skeletonization} algorithm because it only adds $1$-faces (edges) in the inhibitor set $W$ (used for the visual aspect preservation step) (on line \ref{curvskel:01} of Alg.~\ref{algo::3d_curvskel}). For the same reason, we say that the output of Alg.~\ref{algo::3d_curvskel} is a \myem{curvilinear skeleton}.
